---
title: "Judy Li"
author: "FINAL EXAM 3040/5040"
date: "2022-12-14"
output:
  word_document: default
  html_document:
    df_print: paged
subtitle: wl632
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn = -1)

library(stringr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(readr)
library(rvest)
library(patchwork)
```





\newpage





# Problem #1: RMD





\newpage





## Problem #2: HEADERS





\newpage





### Problem #3: Lists

1. Asia
2. Europe
3. North America
    + United States
    + Canada
4. South America
5. Africa





\newpage





#### Problem #4: Fancy Text

*Space: the final frontier.* These are the voyages of the ~starship~ ^Enterprise^. Its five-year mission: to explore strange new worlds. To seek out new life and new civilizations. **To boldly go where no man has gone before!**





\newpage





##### Problem #5: Code Chunks





\newpage





###### Problem #6: Packages





\newpage





# Problem #7: IMPORTING RDATA - LEAVES.RDATA

```{r Q07 }
# Load the LEAVES_3040 dataframe from the LEAVES.RDATA file
load("./GET_LEAVES/LEAVES.RDATA")

# Access the fiftieth row of the data
fiftieth_row <- LEAVES_3040[50,]
fiftieth_row 
```





\newpage





# Problem #8: INLINE CODE - LEAVES_3040 - LEAVES.RDATA

```{r Q08}
# find the length of the leaf on the fiftieth line
fiftieth_row$Length
```





\newpage





# Problem #9: PLOTS A - LEAVES_3040 - LEAVES.RDATA

```{r Q09, echo=FALSE, fig.width=4,fig.height=4}
LEAVES_3040$Year <- factor(LEAVES_3040$Year)
LEAVES_3040 <- mutate(LEAVES_3040, type=ifelse(Year %in% c("2022","2019"), "Highlighted", "Normal"))

# create a boxplot using ggplot
ggplot(LEAVES_3040, aes(x = Length, y = Year, fill=type, group = Year)) +
  geom_boxplot(show.legend = FALSE) +
  scale_fill_manual(values=c("red", "white")) +
  labs(title = "Cornell LEAVES ( 2019 - 2022)",caption = "This is coded to be 4 x 4 inches." , x="Length", y="YEAR")
```





\newpage





# Problem #10: Summary Statistics - LEAVES_3040 - LEAVES.RDATA

```{r Q10 }

# Group the data by shape, quad, and year
grouped_data <- group_by(LEAVES_3040, Shape, Quad, Year)

# Calculate the mean and standard deviation for the Length variable for each group
summary_data <- summarize(grouped_data, mean = mean(Length),sd = sd(Length))

summary_data
```





\newpage





# Problem #11: PLOTS A - LEAVES_3040 - LEAVES.RDATA

```{r Q11, fig.width=5, fig.height=4,echo=FALSE}
stat_args <- group_by(LEAVES_3040, Quad, Year)
stat_args <- summarize(stat_args, stats = c(mean(Length), max(Length), min(Length), median(Length)), Statistic = c('Average', 'Maximum', 'Minimum', 'Median'))

# Plot the histograms with facets
ggplot(LEAVES_3040, aes(Length)) +
  geom_histogram(position="dodge", fill="red") +
  facet_grid(Quad ~ Year) +
  geom_point(aes(x=stats, color = Statistic), y=0, shape = 17, size = 2, data=stat_args) + 
  theme(legend.position="right") +
  labs(title = "Cornell LEAVES ( 2019 - 2022)",caption = "This is coded to be 4 x 5 inches.", x="Length", y="Frequency")

```





\newpage





# Problem #12: PLOTS B - LEAVES_3040 - LEAVES.RDATA

```{r Q12, fig.width=7,fig.height=2,echo=FALSE}
Quad_newdata <- ifelse(LEAVES_3040$Quad == "AG", -0.2, 0.2)

# create the first boxplot
boxplot.1 <- ggplot(data = LEAVES_3040, mapping = aes(x = Quad_newdata, y = Length, fill = Quad))+
  geom_boxplot(width = 0.35)+
  scale_fill_manual(values = c("red","white"))+
  scale_x_continuous(limits = c(-0.4,0.4), breaks = c(-0.2,0.0,0.2), labels = c("-0.2","0.0","0,2"))+
  labs(title = "Boxplot",subtitle = "It is the lengths again.")+
  xlab("")

# create the second histogram
histogram.2 <- ggplot(data = LEAVES_3040, mapping =aes(x= Length, fill= Quad))+
  geom_histogram(position = "identity", alpha = 0.5)+ 
  scale_fill_manual(values = c("red","white"))+
  labs(title = "LEAVES_3040 Histogram",subtitle = "One behind the other")+
  labs(title = "LEAVES_3040 Histogram",fill = "Quad")

# create the third quantile plot
quantile.3 <-ggplot(data = LEAVES_3040, aes(sample= Length, color= Quad)) +
  geom_qq(alpha = 0.2 , distribution = stats::qnorm, dparams = list(mean= 0, sd = 1 ))+
  geom_qq_line(distribution = stats::qnorm, dparams = list( mean= 0, sd = 1))+
  facet_wrap(~LEAVES_3040$Quad)+
  coord_flip()+
  scale_color_manual(values = c("red","black"))+
  labs(title = "Quantile Plot")+
  xlab("theoretical")+
  ylab("sample")
 
# combine three plots into one
boxplot.1 + histogram.2 + quantile.3

```





\newpage





# Problem #13: FUNCTIONS A - LEAVES_3040 - LEAVES.RDATA

```{r Q13 }
problem.13 <- function(df){
  number = 0
  tear_shaped = df$Shape
  # Compute the total number of leaves
  for (i in (1:length(tear_shaped))){
    if (tear_shaped[i] == "T"){ 
      number = number + 1
  } }
  # Return the proportion of tear-shaped leaves
  return ((number)/length(tear_shaped))
}

problem.13(LEAVES_3040)

```





\newpage





# Problem #14: FUNCTIONS B - LEAVES_3040 - LEAVES.RDATA

```{r Q14}
problem.14 <- function(numeric.vector){
length_of_vector = length(numeric.vector)
ordered = sort(numeric.vector, decreasing = FALSE) 
if (length_of_vector %% 2 ==1){ 
  # compute the index of q1,median,q3 for vector with odd number of elements 
  med_index = (length_of_vector + 1)/2  
  q1_index = med_index/2 
  q3_index = med_index + q1_index 
  # compute the five number summary for vector with odd number of elements
  min_val = ordered[1]
  q1_val = (ordered[q1_index] + ordered[q1_index+1])/2
  med_val = ordered[med_index]
  q3_val = (ordered[q3_index] + ordered[q3_index-1])/2
  max_val = ordered[length_of_vector]
} 
else{ 
  #  compute the index of q1,median,q3 for vector with even number of elements 
  med_index = (length_of_vector)/2 
  q1_index = (med_index+1)/2 
  q3_index = med_index + q1_index 
  # compute the five number summary for vector with even number of elements
  min_val = ordered[1]
  q1_val = ordered[q1_index]
  med_val =(ordered[med_index]+ordered[med_index+1])/2
  q3_val = ordered[q3_index]
  max_val = ordered[length_of_vector]
}
  return (c(min_val, q1_val, med_val, q3_val, max_val))
}

# compute the five number summary of the lengths of leaves in LEAVES_3040
problem.14(LEAVES_3040$Length)
```





\newpage





# Problem #15: Combining Data - FOUR YEARS DATA

```{r Q15, echo=FALSE}
base_URL = "http://jeremyentner.com/3040_Final_Data/CompactData/3010_LEAVES"

# define URLs for the files
years = c('2019', '2020', '2021', '2022')
urls <- c("http://jeremyentner.com/3040_Final_Data/CompactData/3010_LEAVES2019.csv", 
          "http://jeremyentner.com/3040_Final_Data/CompactData/3010_LEAVES2020.csv",
          "http://jeremyentner.com/3040_Final_Data/CompactData/3010_LEAVES2021.csv",
          "http://jeremyentner.com/3040_Final_Data/CompactData/3010_LEAVES2022.csv")

# create empty list to store data frames
data_frames <- list()

# loop through URLs and read in each file as a data frame
for (year in years) {
  df <- read_csv(paste(c(base_URL, year, '.csv'), collapse=''), show_col_types = FALSE)
  names(df) <- toupper(names(df))
  if (year == '2019') { 
    df <- mutate(df, SHAPE = as.character(T_NT))  
  }
  df <- mutate(df, STUDENT = as.character(STUDENT), YEAR = year)
  data_frames[[year]] <- df
}

# combine data frames into a single data frame
year_total <- c()
shape_total <- c()

for (i in (1:length(urls))) {
  year_total <- append(year_total, data_frames[[i]]$YEAR)
  shape_total <- append(shape_total, data_frames[[i]]$SHAPE)
}

combined_df <- data.frame(YEAR=year_total, SHAPE=shape_total)
combined_df <- mutate(combined_df, SHAPE = ifelse(toupper(SHAPE) == 'T', "Tear", "Other"))
combined_df <- drop_na(combined_df, YEAR)
combined_df <- group_by(combined_df, YEAR, SHAPE)
combined_df <- summarize(combined_df, total = n())

# create stacked bar chart
ggplot(combined_df, aes(x = YEAR, y = total, fill = SHAPE)) +
  geom_bar(position="fill", stat="identity", color="black") + 
  labs(title = "CORNELL LEAVES", y = "Proportion", x = "YEAR") +
  scale_fill_manual(values = c("white", "red"), breaks=c('Tear', 'Other'))

```





\newpage





# Problem #16: SCRAPING - PART 1

```{r Q16, echo=FALSE}
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}

quadHandle <- function(x){
  if_else(grepl("AG", x, fixed=TRUE), 'AG', "ART")
}

base_URL = "http://jeremyentner.com/3040_Final_Data/"

# create empty list to store data frames
data_frames <- list()

scrape_hrefs <- function(url) {
  # Read the HTML from the URL
  html <- read_html(url)
  
  # Use CSS selectors to find all `a` tags on the page
  a_tags <- html_nodes(html, "a")
  
  # Extract the `href` attribute from each `a` tag
  hrefs <- html_attr(a_tags, "href")
  
  return(hrefs)
}

hrefs <- scrape_hrefs("http://jeremyentner.com/3040_Final_Data/HtmlExampleScraping1.html")

# loop through URLs and read in each file as a data frame
for (url in hrefs) {
  first_year = substring(url, 0, 4)
  last_year = sub(".csv", "", substrRight(url, 8))
  if(startsWith(url, '2021') || startsWith(url, '2022')){
    df <- read_csv(paste(c(base_URL, url), collapse=''), col_names = c('length','shape','date'), show_col_types = FALSE)
    quad <- sub(".CSV", "", strsplit(url, "/")[[1]][2])
    section <- sub(".CSV", "", strsplit(strsplit(url, "/")[[1]][2], "_")[[1]][2])
    df <- mutate(df, year = first_year, quad = quad, section = section)
  } else {
    df <- read_csv(paste(c(base_URL, url), collapse=''), show_col_types = FALSE)
    names(df) <- gsub(" ", "_", names(df))
    df <- mutate(df, year = last_year)
  }
  
  names(df) <- toupper(names(df))
  if("LAB_SECTION" %in% names(df)){
      df <- mutate(df, SECTION = LAB_SECTION)
  }
  
  if (last_year == '2019') { 
    df <- mutate(df, SHAPE = as.character(T_NT), LENGTH = LENGTHS, QUAD = AG_ART)  
  }
  
  df <- mutate(df, QUAD = quadHandle(QUAD), SECTION = as.factor(SECTION))
  data_frames[[url]] <- df
}

# combine data frames into a single data frame
length_total <- c()
year_total <- c()
section_total <- c()
quad_total <- c()

for (i in (1:length(hrefs))) {
  length_total <- append(length_total, data_frames[[i]]$LENGTH)
  year_total <- append(year_total, data_frames[[i]]$YEAR)
  section_total <- append(section_total, data_frames[[i]]$SECTION)
  quad_total <- append(quad_total, data_frames[[i]]$QUAD)
}

combined_df <- data.frame(LENGTH=length_total, YEAR=year_total, SECTION=section_total, QUAD=quad_total)
combined_df <- mutate(combined_df, SECTION = as.numeric(as.character(SECTION)))
combined_df <- combined_df[order(combined_df$SECTION, decreasing = TRUE),]
combined_df <- mutate(combined_df, SECTION = as.factor(SECTION))

# Plot the scatterplot with facets
ggplot(combined_df, aes(x=YEAR, y=LENGTH)) +
  geom_jitter(aes(colour = SECTION),size=0.3) +
  facet_grid(~QUAD) +
  theme(legend.position="right") +
  labs(title = "CORNELL LEAVES - NOT CORNELL COLORS", x="YEAR", y="LENGTH")

```





\newpage





# Problem #17: SCRAPING - PART 2

```{r Q17, echo=FALSE}
quadHandle <- function(x){
  if_else(grepl("AG", x, fixed=TRUE), 'AG', "ART")
}

base_URL = "http://jeremyentner.com/3040_Final_Data/"

# create empty list to store data frames
data_frames <- list()

hrefs <- scrape_hrefs("http://jeremyentner.com/3040_Final_Data/HtmlExampleScraping2.html")


# loop through URLs and read in each file as a data frame
for (url in hrefs) {
  first_year = substring(url, 0, 4)
  last_year = sub(".csv", "", substrRight(url, 8))
  if(startsWith(url, '2021') || startsWith(url, '2022')){

    df <- read_csv(paste(c(base_URL, url), collapse=''), show_col_types = FALSE)
    
    if(length(names(df)) > 2) {
      if ("...1" %in% names(df) || "Circle Number" %in% names(df)){
        df <- select(df, 2:4)
      }
      df <- select(df, 1:3)
      names(df) <- c('length','shape','date')
    }
    else {
      names(df) <- c('length','shape')
    }
  
    quad <- sub(".CSV", "", strsplit(url, "/")[[1]][2])
    section <- sub(".CSV", "", strsplit(strsplit(url, "/")[[1]][2], "_")[[1]][2])
    df <- mutate(df, year = first_year, quad = quad, section = section, length = as.numeric(gsub("[^0-9.-]", "", length)))
  } else {
    df <- read_csv(paste(c(base_URL, url), collapse=''), show_col_types = FALSE)
    df <- mutate(df, year = last_year)
  }
  
  names(df) <- gsub(" ", "_", toupper(names(df)))

  if("LAB_SECTION" %in% names(df)){
      df <- mutate(df, SECTION = LAB_SECTION)
  }
  
  if (last_year == '2019') { 
    df <- mutate(df, SHAPE = as.character(T_NT), LENGTH = LENGTHS, QUAD = AG_ART)  
  }
  
  df <- mutate(df, QUAD = quadHandle(QUAD), SECTION = as.factor(SECTION), SHAPE = as.factor(SHAPE))
  data_frames[[url]] <- df
}

# combine data frames into a single data frame
length_total <- c()
year_total <- c()
section_total <- c()
shape_total <- c()

for (i in (1:length(hrefs))) {
  length_total <- append(length_total, data_frames[[i]]$LENGTH)
  year_total <- append(year_total, data_frames[[i]]$YEAR)
  section_total <- append(section_total, data_frames[[i]]$SECTION)
  shape_total <- append(shape_total, data_frames[[i]]$SHAPE)
}

combined_df <- data.frame(LENGTH=length_total, YEAR=year_total, SECTION=section_total, SHAPE=shape_total)
combined_df <- mutate(combined_df, SHAPE = toupper(SHAPE), SECTION = as.numeric(as.character(SECTION)))
combined_df <- drop_na(combined_df, LENGTH, SECTION)
combined_df <- combined_df[order(combined_df$SECTION, decreasing = TRUE),]
combined_df <- mutate(combined_df, SHAPE = ifelse(toupper(SHAPE) == 'T', "Tear", "Other"))

# Plot the bar chart with facets
ggplot(combined_df, aes(x = LENGTH, fill = SHAPE)) + 
  geom_histogram(color="black") +
  facet_wrap(~ SECTION) + 
  labs(title = "Cornell LEAVES (2019-2022)", x="Length", y="Frequency") + 
  scale_fill_manual(values = c("white", " red"), breaks=c('Tear', 'Other'))
  
```





\newpage





# Problem #18 - Submission

I submitted it to GRADESCOPE. I submitted it to CANVAS.
